A Mechanised Library for Generic (Co)Programming

### 2024-02-06 paper/paper.tex
Updates on containers

\section{Mechanising Extractable Containers}

@@

Our goals are: (1) the mechanisation of hylomorphisms and their laws should be
axiom-free; and (2) Coq hylomorphisms should be extracted to recursive OCaml
functions that use no instances of \mintinline{OCaml}{Obj.magic}. Note that
although the use of \mintinline{OCaml}{Obj.magic} is safe in the code extracted
by Coq, it increases the Trusted Code Base, makes code less legible for
programmers, and its presence will likely inhibit optimisations by the OCaml
compiler\footnote{\dcas{I read this somewhere, search \& add reference/link?}}.

@@

Mechanising functors in Coq is straightforward. But taking least/greatest
fixpoints of the said functors may fail, due to not being strictly positive. A
common approach to construct strictly positive functors in Coq is to use
\emph{containers}. In essence, a container 
\dcas{A first approach (problems with extraction, lots of Obj.magic, etc)}
\begin{minted}{Coq}
  Record Container (Shape : Set) (Pos : Shape -> Set) (X : Set) :=
    { shape : Shape; contents : Pos shape -> X }.
\end{minted}

>>>>

\section{Mechanising Extractable Container Functors}

@@

Our mechanisation works in the category of Coq types, and our goals are: (1)
the mechanisation of hylomorphisms and their laws should be axiom-free; and (2)
Coq hylomorphisms should be extracted to recursive OCaml functions that use no
instances of \mintinline{OCaml}{Obj.magic}.  Although the use of
\mintinline{OCaml}{Obj.magic} is safe in the code extracted by Coq, it has been
proven that, for higher-order programs, simple interoperations can lead to
incorrect behaviour or even segfaults~\cite{forster:hal-04329663}.  Therefore,
the avoidance of axioms and  \mintinline{OCaml}{Obj.magic} means that the
extracted code will have stronger safety guarantees, and will allow the safe
interoperation with other OCaml code.

@@

In Coq, to construct least/greatest fixpoints of functors, these need to build
\emph{strictly positive types}. A common approach to construct such strictly
positive endofunctors is to use \emph{container functors}~\cite{Whatever}.
However, reasoning about container and container functor equality requires to
consider: (a) functional extensionality; (b) heterogeneous equality. This is
problematic, due to our goals to remain axiom-free. Furthermore, standard
representations of containers in Coq will lead to OCaml code that contains
\mintinline{OCaml}{Obj.magic}.

A containers is defined by a type of \emph{shapes} (\mintinline{Coq}{Shape }),
and a \emph{family of position types} indexed by shapes
(\mintinline{Coq}{Pos}).  An \emph{extension} of this container is a functor,
whose action on objects is given by the following definition:
\begin{minted}{Coq}
  Record Container (Shape : Set) (Pos : Shape -> Set) (X : Set) :=
    { shape : Shape; contents : Pos shape -> X }.
\end{minted}
Note that, in our mechanisation, we call container extensions
\mintinline{Coq}{Container}. Generally, any functor that is isomorphic to a
container extension is called a \emph{container functor}. However, in our work
we will call \emph{container functor} to the functors defined by a specific
container extension.

The first problem in extracting Coq code from such container extensions is
the use of families of types. Since OCaml's
